---
title: "Pension Plans vs. Benchmarks"
subtitle: "MSCI ACWI ex-U.S stocks, Russell 3000 ETF (IWV), Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)"
date: January 15, 2021
author: "Truong Bui, Thuy Nguyen" 
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(quantmod)
library(PerformanceAnalytics)
library(quadprog)
library(knitr)
library(kableExtra)

my_scroll_table <- function(df){

  kable(df, "html") %>%
  kable_styling() %>%
  scroll_box(height = "500px", width = "800px")
}
```


# Data 
## Benchmarks
[Richard M. Ennis (2020)](https://richardmennis.com/blog/endowment-performance) used three public market indexes: 

* Bloomberg Barclays U.S. Aggregate bonds, 
* Russell 3000 stocks 
* MSCI ACWI ex-U.S stocks.

Except for the MSCI ACWI ex-U.S stocks, which is retrieved from [MSCI website](https://www.msci.com/end-of-day-data-regional), we use 2 other benchmarks for substitution:  

* [iShares Russell 3000 ETF (IWV)](https://www.ftserussell.com/products/indices/russell-us)

* [Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)](https://finance.yahoo.com/quote/VBTIX?p=VBTIX&.tsrc=fin-srch) 

*Note: iShares Russell 3000 ETF (IWV) and Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX) use the first day of the month*

### acwi_exUS_price
```{r acwi_exUS_price}
# https://www.msci.com/end-of-day-data-regional
acwi_exUS <- rio::import(here::here("data", "historyIndex.xls"), skip=6)

acwi_exUS_price <- acwi_exUS %>% slice(1:252) %>% # last few rows are not numbers
  rename("acwi_exUS" = `ACWI ex USA Standard (Large+Mid Cap)`) %>% 
  mutate(acwi_exUS = as.numeric(acwi_exUS)) %>% 
  mutate(Date = str_remove(Date, ",")) %>% 
  mutate(Date = as.Date(Date, "%B%d%Y"),
         month = month(Date), 
         fy = year(Date)) %>% 
  filter(Date >= "2000-06-01") %>% # pension data only has from 2000-06-01
  select(-Date) 

head(acwi_exUS_price)
```
### russell3000_IWV_price
```{r russell3000_IWV_price}
#iShares Russell 3000 ETF (IWV)
#https://finance.yahoo.com/quote/IWV/history?p=IWV
russell3000_IWV <- rio::import(here::here("data", "IWV.csv"))

russell3000_IWV_price <- russell3000_IWV %>% 
    select(Date, `Adj Close`) %>% # get the Adj Close price
    rename("russell3000_IWV" = `Adj Close`) %>% 
    mutate(month = month(Date), 
           fy = year(Date)) %>% 
    filter(Date < "2021-01-01")

head(russell3000_IWV_price)
```

### vanguardBond_VBTIX_price
```{r vanguardBond_VBTIX_price}
#Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)

vanguardBond_VBTIX <- rio::import(here::here("data", "VBTIX.csv"))

vanguardBond_VBTIX_price <- vanguardBond_VBTIX %>% 
    select(Date, `Adj Close`) %>% 
  filter(Date >= "2000-06-01") %>% # pension data only has from 2000-06-01
  rename("vanguardBond_VBTIX" = `Adj Close`) %>% 
  mutate(month = month(Date), 
         fy = year(Date)) %>% 
  filter(Date < "2021-01-01")

head(vanguardBond_VBTIX_price)
```

## Pension data
```{r}
d <- rio::import(here::here("data", "ppd-data-latest-7.csv"))

pension_data <- d %>% 
  filter(fy >= 2000) %>%
  select(PlanName, PlanFullName, fy, fye, ActLiabilities_GASB, InvestmentReturn_1yr) %>% # Filter some key variables
  mutate(PlanFullName = gsub("\x92", "'", PlanFullName), # Fix plans' full names
         PlanName = gsub("\x92", "'", PlanName)) %>% 
  mutate(fye = ymd(fye),
         month = month(fye))

my_scroll_table(head(pension_data))
```

## Joining data

```{r price_year}
index_price <- left_join(vanguardBond_VBTIX_price, russell3000_IWV_price) %>% 
  left_join(acwi_exUS_price) %>% 
  select(Date, month, fy, vanguardBond_VBTIX, russell3000_IWV, acwi_exUS)

# only keep data of the months that Pension data has
pension_month <- unique(pension_data$month)

price_year <- index_price %>% 
  filter(month %in% pension_month) 

my_scroll_table(head(price_year))
```


```{r stock_returns}
# Calculate annual returns for the individual securities
stock_returns <- price_year %>%
  data.frame(date = index(.)) %>%
  pivot_longer(cols = 4:6, 
               names_to = "index", 
               values_to = "price") %>%
  arrange(index, month, fy) %>% 
  group_by(index, month) %>%
  mutate(returns = price/lag(price) - 1) %>%
  select(-price, -date) %>%
  pivot_wider(names_from = index, values_from = returns) %>% 
  ungroup()

stock_returns %>% head()
```

```{r pension_index}
#Join pension data with stock returns
pension_index <- pension_data %>% 
  left_join(stock_returns, by = c("fy", "month")) %>% 
  select(-c(PlanFullName, fy, month, ActLiabilities_GASB)) %>% drop_na()

#write.csv(pension_index, "pension_index.csv")

my_scroll_table(head(pension_index))
```

# Testing on one case: Alabama ERS
```{r Alabama ERS}
alabama <- pension_index %>% 
  filter(PlanName == "Alabama ERS") 
head(alabama)
```
## Optimizing Portfolio Weights 


```{r optimize, best fit}
#document of the package here: https://cran.r-project.org/web/packages/quadprog/quadprog.pdf
# other package Dyktra 2018: https://cran.r-project.org/web/packages/Dykstra/Dykstra.pdf
# other: pracma, using lsqlincon: https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/lsqlincon
y_actual = matrix(alabama$InvestmentReturn_1yr)

x1 = matrix(alabama$acwi_exUS)
x2 = matrix(alabama$russell3000_IWV)
x3 = matrix(alabama$vanguardBond_VBTIX)
x = cbind(x1, x2, x3)

#method 1: 
# Dmat <- t(x) %*% x 
# Amat <- cbind(rep(1,3), diag(3)) # vector define constraint
# bvec <- c(1,rep(0,3)) #vector of b coefficient
# dvec <- t(y_actual) %*% x  # vector to be minimized: product of transpose y and x
# result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)
# result

#method 2: more elegant using R base function crossprod
Dmat <- crossprod(x)
dvec <- crossprod(y_actual, x)  # vector to be minimized: product:transpose y_actual and x
Amat <- cbind(rep(1,3), diag(3)) # vector define constraint
bvec <- c(1, 0, 0, 0) #vector of b coefficient; meq = 1 is equality constraint: coefs sum to 1

result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1) 
result
```
## Apply the best fit solution to Alabama

```{r}
#apply the best fit result to alabama
alabama_portfolio_result <- alabama %>% 
  add_column(b1_acwi_exUS = result$solution[1]) %>% 
  add_column(b2_russell3000_IWV = result$solution[2]) %>% 
  add_column(b3_vanguardBond_VBTIX = result$solution[3]) %>% 
  mutate(portfolio_return = acwi_exUS*b1_acwi_exUS + russell3000_IWV*b2_russell3000_IWV + 
           vanguardBond_VBTIX*b3_vanguardBond_VBTIX) %>% 
  
  # sum of square total: (actual value - average of actual value)^2
  mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 

  # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
  mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 

  # r_squared
  mutate(r_squared = 1 - (sse/sst))

my_scroll_table(head(alabama_portfolio_result))
```

# Writing functions to generalize 

## optimize_beta 
```{r optimize_beta function}

optimize_beta <- function(planName){

  df <- pension_index %>% 
  filter(PlanName == planName) 
  
y_actual = matrix(df$InvestmentReturn_1yr)
x1 = matrix(df$acwi_exUS)
x2 = matrix(df$russell3000_IWV)
x3 = matrix(df$vanguardBond_VBTIX)
x = cbind(x1, x2, x3)

Dmat <- crossprod(x)
dvec <- crossprod(y_actual, x)
Amat <- cbind(rep(1,3), diag(3)) 
bvec <- c(1, 0, 0, 0) 
dvec <- t(y_actual) %*% x  
result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)

return(result)
  
}

optimize_beta("Alabama ERS")

```

## apply_portfolio_to_plan
```{r apply_portfolio_to_plan function}

apply_portfolio_to_plan <- function (planName) {
  
  plan <- pension_index %>% 
  filter(PlanName == planName) 
  
  result <- optimize_beta(planName)
  
  portfolio_result <- plan %>% 
  add_column(b1_acwi_exUS = result$solution[1]) %>% 
  add_column(b2_russell3000_IWV = result$solution[2]) %>% 
  add_column(b3_vanguardBond_VBTIX = result$solution[3]) %>% 
  mutate(portfolio_return = acwi_exUS*b1_acwi_exUS + russell3000_IWV*b2_russell3000_IWV + 
           vanguardBond_VBTIX*b3_vanguardBond_VBTIX) %>% 
  
  # sum of square total: (actual value - average of actual value)^2
  mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 

  # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
  mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 

  # r_squared
  mutate(r_squared = 1 - (sse/sst))

 return(portfolio_result)
  
}

apply_portfolio_to_plan("Alabama ERS") %>% head()
```

## geo_mean

```{r}
#Can't believe R does not have a built-in function for this.
geo_mean <- function(vector) {
  vector <- vector + 1
  r <- prod(vector)^(1/length(vector)) - 1
  return(r)
}
```

## comparison_result 

```{r comparison_result function}

comparison_plan_benchmark <- function(df) {
  
comparison_result <- df %>% 
  
  #geomean
  mutate(geomean_actual = geo_mean(InvestmentReturn_1yr)) %>% 
  mutate(geomean_portfolio = geo_mean(portfolio_return)) %>% 
  
  # standard deviation (i.e. risk)
  mutate(sd_actual = sd(InvestmentReturn_1yr)) %>% 
  mutate(sd_portfolio = sd(portfolio_return))

return(comparison_result)
}

my_scroll_table(head(comparison_plan_benchmark(apply_portfolio_to_plan("Alabama ERS"))))
```

# Applying the function - Case testing

Applying the function to examine 2 individual pension plans. 

## California PERF
```{r California}
my_scroll_table(head(comparison_plan_benchmark(apply_portfolio_to_plan("California PERF"))))
```
## South Dakota RS
```{r}
my_scroll_table(head(comparison_plan_benchmark(apply_portfolio_to_plan("South Dakota RS"))))
```

# Applying the function to all pension plans

## Pension plans got processed
```{r}
# Get a list of the first 5 plans
plan_list <- (unique(pension_index$PlanName))

# Loop through each plan to produce result
all_plans = list()

for (plan in plan_list) {
  skip_to_next <- FALSE # initialize FALSE
  
  tryCatch(
  result <- comparison_plan_benchmark(apply_portfolio_to_plan(plan)),
  
  error = function(e) { 
    skip_to_next <<- TRUE} # if get an error, trigger to TRUE 
  )
  
  all_plans[[plan]] <- result
  if(skip_to_next) { next } # just skip that item and go to the next
}

final_result <- bind_rows(all_plans)  

my_scroll_table(head(final_result))
write_csv(final_result, "final_result.csv")
```
## Pension plans do not get processed

The subset of any left over that the function above does not process. 


```{r}
error_plans <- setdiff(plan_list, unique(final_result$PlanName))
error_plans

```



