---
title: "Pension Plans vs. Benchmarks - 10 Years"
subtitle: "MSCI ACWI ex-U.S stocks, Russell 3000 ETF (IWV), Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)"
date: May 15, 2021
author: "Truong Bui, Thuy Nguyen" 
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(quantmod)
library(PerformanceAnalytics)
library(quadprog)
library(knitr)
library(kableExtra)
library(stringr)

my_scroll_table <- function(df){

  kable(df, "html") %>%
  kable_styling() %>%
  scroll_box(height = "500px", width = "800px")
}

## geo_mean
#Can't believe R does not have a built-in function for this.
geo_mean <- function(vector) {
  vector <- vector + 1
  r <- prod(vector)^(1/length(vector)) - 1
  return(r)
}
```



# Data 
## Pension
```{r}
d2 <- rio::import(here::here("data", "ppd-data-latest_may2022.csv")) # Following Ennis, Download Full dataset here https://publicplansdata.org/public-plans-database/download-full-data-set/#nav-1

# Pension data 10yr investment return in 2020
pension_data2 <- d2 %>% 
  filter(fy > 2010 & fy < 2021) %>%
  select(PlanName, PlanFullName, fy, fye, InvestmentReturn_1yr, InvestmentReturn_10yr) %>% # Filter some key variables
  mutate(PlanFullName = gsub("\x92", "'", PlanFullName), # Fix plans' full names
         PlanName = gsub("\x92", "'", PlanName)) %>% 
  mutate(fye = ymd(fye),
         month = month(fye)) 
```


## Benchmarks
[Richard M. Ennis (2020)](https://richardmennis.com/blog/endowment-performance) used three public market indexes: 

* Bloomberg Barclays U.S. Aggregate bonds, 
* Russell 3000 stocks index
* MSCI ACWI ex-U.S stocks.

Except for the MSCI ACWI ex-U.S stocks, which is retrieved from [MSCI website](https://www.msci.com/end-of-day-data-regional), we use 2 other benchmarks for substitution:  

* [iShares Russell 3000 ETF (IWV)](https://www.ftserussell.com/products/indices/russell-us)

* [Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)](https://finance.yahoo.com/quote/VBTIX?p=VBTIX&.tsrc=fin-srch) 

*Note: iShares Russell 3000 ETF (IWV) and Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX) use the first day of the month*

### MSCI acwi_exUS_price
```{r acwi_exUS_price}
# https://www.msci.com/end-of-day-data-regional
acwi_exUS <- rio::import(here::here("data", "historyIndex_may2022.xls"), skip=6) 

acwi_exUS_price <- acwi_exUS %>%  
slice(1:264) %>% # last few rows are not numbers
  rename("acwi_exUS" = `ACWI ex USA Standard (Large+Mid Cap)`) %>% 
  mutate(acwi_exUS = as.numeric(acwi_exUS)) %>% 
  mutate(Date = str_remove(Date, ",")) %>% 
  mutate(Date = as.Date(Date, "%B%d%Y"),
         month = month(Date), 
         fy = year(Date)) %>% 
  filter(Date <= "2021-06-30") %>% 
  select(-Date) 

head(acwi_exUS_price)
```


### russell3000_IWV_price
```{r russell3000_IWV_price}
#iShares Russell 3000 ETF (IWV)
#https://finance.yahoo.com/quote/IWV/history?p=IWV
russell3000_IWV <- rio::import(here::here("data", "IWV.csv"))

russell3000_IWV_price <- russell3000_IWV %>% 
    select(Date, `Adj Close`) %>% # get the Adj Close price
    rename("russell3000_IWV" = `Adj Close`) %>% 
    mutate(month = month(Date), 
           fy = year(Date)) %>% 
    filter(Date <= "2021-06-30")

head(russell3000_IWV_price)
```

### vanguardBond_VBTIX_price
```{r vanguardBond_VBTIX_price}
#Vanguard Total Bond Market Index Fund Institutional Shares (VBTIX)

vanguardBond_VBTIX <- rio::import(here::here("data", "VBTIX.csv"))

vanguardBond_VBTIX_price <- vanguardBond_VBTIX %>% 
    select(Date, `Adj Close`) %>% 
  filter(Date >= "2000-06-01") %>% # pension data only has from 2000-06-01
  rename("vanguardBond_VBTIX" = `Adj Close`) %>% 
  mutate(month = month(Date), 
         fy = year(Date)) %>% 
  filter(Date <= "2021-06-30")

head(vanguardBond_VBTIX_price)
```


## Joining data

```{r price_year}
index_price <- left_join(vanguardBond_VBTIX_price, russell3000_IWV_price) %>% 
  left_join(acwi_exUS_price) %>% 
  select(Date, month, fy, vanguardBond_VBTIX, russell3000_IWV, acwi_exUS)

# only keep data of the months that Pension data has
pension_month <- unique(pension_data2$month)

price_year <- index_price %>% 
  filter(month %in% pension_month) 

my_scroll_table(head(price_year))
```


```{r stock_returns}
# Calculate annual returns for the individual securities
stock_returns <- price_year %>%
  data.frame(date = index(.)) %>%
  pivot_longer(cols = 4:6, 
               names_to = "index", 
               values_to = "price") %>%
  arrange(index, month, fy) %>% 
  group_by(index, month) %>%
  mutate(returns = price/lag(price) - 1) %>%
  select(-price, -date) %>%
  pivot_wider(names_from = index, values_from = returns) %>% 
  ungroup()

stock_returns %>% head()
```

```{r pension_index}
#Join pension data with stock returns
pension_index <- pension_data2 %>% 
  left_join(stock_returns, by = c("fy", "month")) %>% 
  select(-c(PlanFullName, fy, month)) %>% drop_na()

#write.csv(pension_index, "pension_index.csv")

my_scroll_table(head(pension_index))
```

### Checking Plans' 10 years return against Ennis's Report 

```{r Arizona SRS}

arizona <- pension_index %>% 
   filter(PlanName == "Arizona SRS") 
 geo_mean(arizona$InvestmentReturn_1yr) # Yes, matched
```


```{r New York State Teachers}
newyork_state <- pension_index %>% 
  filter(PlanName == "New York State Teachers") 
geo_mean(newyork_state$InvestmentReturn_1yr) # Yes, matched

```


```{r}
pension_index %>% filter(str_detect(PlanName, "Vermont"))

vermont <- pension_index %>% 
  filter(PlanName == "Vermont Teachers") 

geo_mean(vermont$InvestmentReturn_1yr) # NOT matched
```


```{r}
pension_index %>% filter(str_detect(PlanName, "Texas"))

texas <- pension_index %>%
filter(PlanName == "Texas Teachers") # correct
geo_mean(texas$InvestmentReturn_1yr)# correct

```

```{r}
pension_index %>% filter(str_detect(PlanName, "Ohio"))
ohio <- pension_index %>% 
  filter(PlanName == "Ohio School Employees")
geo_mean(ohio$InvestmentReturn_1yr) # not matched with Ennis's report
```

```{r}
pension_index %>% filter(str_detect(PlanName, "Missouri"))
missouri <- pension_index %>% 
  filter(PlanName == "Missouri State Employees")
geo_mean(missouri$InvestmentReturn_1yr) # Yes, matched
```

## Optimizing Portfolio Weights 


```{r optimize, best fit}
#document of the package here: https://cran.r-project.org/web/packages/quadprog/quadprog.pdf
# other package Dyktra 2018: https://cran.r-project.org/web/packages/Dykstra/Dykstra.pdf
# other: pracma, using lsqlincon: https://www.rdocumentation.org/packages/pracma/versions/1.9.9/topics/lsqlincon
y_actual = matrix(arizona$InvestmentReturn_1yr)

x1 = matrix(arizona$acwi_exUS)
x2 = matrix(arizona$russell3000_IWV)
x3 = matrix(arizona$vanguardBond_VBTIX)
x = cbind(x1, x2, x3)

#method 1: 
# Dmat <- t(x) %*% x 
# Amat <- cbind(rep(1,3), diag(3)) # vector define constraint
# bvec <- c(1,rep(0,3)) #vector of b coefficient
# dvec <- t(y_actual) %*% x  # vector to be minimized: product of transpose y and x
# result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)
# result

#method 2: more elegant using R base function crossprod
Dmat <- crossprod(x)
dvec <- crossprod(y_actual, x)  # vector to be minimized: product:transpose y_actual and x
Amat <- cbind(rep(1,3), diag(3)) # vector define constraint
bvec <- c(1, 0, 0, 0) #vector of b coefficient; meq = 1 is equality constraint: coefs sum to 1

result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1) 
result
```
## Apply the best fit solution to Arizona

```{r}
#apply the best fit result to alabama
arizona_portfolio_result <- arizona %>% 
  add_column(b1_acwi_exUS = result$solution[1]) %>% 
  add_column(b2_russell3000_IWV = result$solution[2]) %>% 
  add_column(b3_vanguardBond_VBTIX = result$solution[3]) %>% 
  mutate(portfolio_return = acwi_exUS*b1_acwi_exUS + russell3000_IWV*b2_russell3000_IWV + 
           vanguardBond_VBTIX*b3_vanguardBond_VBTIX) %>% 
  
  # sum of square total: (actual value - average of actual value)^2
  mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 

  # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
  mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 

  # r_squared
  mutate(r_squared = 1 - (sse/sst))

my_scroll_table(head(arizona_portfolio_result))
```

# Writing functions to generalize 

## optimize_beta 
```{r optimize_beta function}

optimize_beta <- function(planName){

  df <- pension_index %>% 
  filter(PlanName == planName) 
  
y_actual = matrix(df$InvestmentReturn_1yr)
x1 = matrix(df$acwi_exUS)
x2 = matrix(df$russell3000_IWV)
x3 = matrix(df$vanguardBond_VBTIX)
x = cbind(x1, x2, x3)

Dmat <- crossprod(x)
dvec <- crossprod(y_actual, x)
Amat <- cbind(rep(1,3), diag(3)) 
bvec <- c(1, 0, 0, 0) 
dvec <- t(y_actual) %*% x  
result <- solve.QP(Dmat = Dmat, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)

return(result)
  
}

optimize_beta("Arizona SRS")

```

## apply_portfolio_to_plan
```{r apply_portfolio_to_plan function}

apply_portfolio_to_plan <- function (planName) {
  
  plan <- pension_index %>% 
  filter(PlanName == planName) 
  
  result <- optimize_beta(planName)
  
  portfolio_result <- plan %>% 
  add_column(b1_acwi_exUS = result$solution[1]) %>% 
  add_column(b2_russell3000_IWV = result$solution[2]) %>% 
  add_column(b3_vanguardBond_VBTIX = result$solution[3]) %>% 
  mutate(portfolio_return = acwi_exUS*b1_acwi_exUS + russell3000_IWV*b2_russell3000_IWV + 
           vanguardBond_VBTIX*b3_vanguardBond_VBTIX) %>% 
  
  # sum of square total: (actual value - average of actual value)^2
  mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 

  # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
  mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 

  # r_squared
  mutate(r_squared = 1 - (sse/sst))

 return(portfolio_result)
  
}

apply_portfolio_to_plan("Arizona SRS") %>% head()
```


## comparison_result 

```{r comparison_result function}

comparison_plan_benchmark <- function(df) {
  
comparison_result <- df %>% 
  
  #geomean
  mutate(geomean_actual = geo_mean(InvestmentReturn_1yr)) %>% 
  mutate(geomean_portfolio = geo_mean(portfolio_return)) %>% 
  
  # standard deviation (i.e. risk)
  mutate(sd_actual = sd(InvestmentReturn_1yr)) %>% 
  mutate(sd_portfolio = sd(portfolio_return))

return(comparison_result)
}

#my_scroll_table(head(comparison_plan_benchmark(apply_portfolio_to_plan("Arizona SRS"))))
```

## Checking against Exhibit 5 in Ennis's report
```{r}
comparison_plan_benchmark(apply_portfolio_to_plan("Arizona SRS")) %>% select(PlanName, geomean_actual, geomean_portfolio,r_squared)
```

# Applying the function - Case testing

Applying the function to examine individual pension plans. 
```{r}
pension_index %>% 
  filter(str_detect(PlanName, "Washington"))
```



```{r}

# Note: NOT correct 10 yr return: "South Dakota", "Ohio School Employees", "Vermont Teachers"

plan_ennis <- c("Arizona SRS", "California PERF",
                "Florida RS","Illinois SERS", "Illinois Teachers", "Iowa PERS", "Maine State and Teacher", "Minnesota GERF",
                "Missouri State Employees", "New Jersey PERS", "New Mexico PERA", "New York State Teachers", "North Carolina Local Government",
                "Ohio School Employees", "Oregon PERS", "Pennsylvania School Employees", "Rhode Island ERS", "South Carolina RS",
                "South Dakota", "Virginia RS", 
                "Texas Teachers", "Vermont Teachers", "Washington PERS 2/3")

for (plan in plan_ennis) {
  result <- comparison_plan_benchmark(apply_portfolio_to_plan(plan)) %>% 
    select(PlanName, geomean_actual, geomean_portfolio,r_squared)
  print(result)
}

```




