---
title: "Pension Plans & Benchmarks"
date: December 8, 2021
author: "Truong Bui, Thuy Nguyen" 
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(quantmod)
library(PerformanceAnalytics)
library(quadprog)
```

# Data Manipulation

## Pension Data

```{r pension data}
# Import pension data
d <- read.csv("ppd-data-latest-7.csv")

pension_data <- d %>% 
  filter(fy >= 2000) %>%
  select(PlanName, PlanFullName, fy, fye, ActLiabilities_GASB, InvestmentReturn_1yr) %>% # Filter some key variables
  mutate(PlanFullName = gsub("\x92", "'", PlanFullName), # Fix plans' full names
         PlanName = gsub("\x92", "'", PlanName)) %>% 
  mutate(fye = ymd(fye),
         month = month(fye))
head(pension_data)
```

## Benchmark Data
```{r index prices}
#Get index prices
symbols <- c("SWTSX", "VBMFX")

prices <- getSymbols(symbols,
                     src = "yahoo",
                     from = "1998-12-31",
                     to = "2020-12-31",
                     auto.assign = T,
                     warnings = F) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>% 
  `colnames<-`(symbols)

#Get yearly prices for months available in pension data only
prices_monthly <- to.monthly(prices, OHLC = F)
unique_month_pension <- unique((pension_data$month))
unique_month <- unique_month_pension[!is.na(unique_month_pension)]
prices_yearly <- prices_monthly[month(index(prices_monthly)) %in% unique_month]

head(prices_yearly)
```

## Merging Pension & Benchmark Data

```{r pension_index data}
# Calculate annual returns for the individual securities
stock_returns <- prices_yearly %>%
  data.frame(date = index(.)) %>%
  remove_rownames() %>%
  mutate(fy = year(date), month = month(date)) %>% 
  pivot_longer(cols = 1:2, names_to = "security", values_to = "prices") %>%
  arrange(security, month, fy) %>% 
  group_by(security, month) %>%
  mutate(returns = prices/lag(prices) - 1) %>%
  select(-prices, -date) %>%
  pivot_wider(names_from = security, values_from = returns) %>% 
  ungroup()

#Join pension data with stock returns
pension_index <- pension_data %>% 
  left_join(stock_returns, by = c("fy", "month")) %>% 
  select(-c(PlanFullName, fy, month, ActLiabilities_GASB)) 

#write.csv(pension_index, "pension_index.csv")
head(pension_index)
```

# Alabama Plan

```{r Alabama ERS}
alabama <- pension_index %>% 
  filter(PlanName == "Alabama ERS") 
head(alabama)
```

## Optimizing Portfolio Weights 

Genetic linear model model with constraint imposed on coefficients: y ~ b1 * X1 + b2 * X2

Subject to contraints: b1 + b2 = 1; b1 >= 0 AND b2 > 0

Translated into our problem as: InvestmentReturn_1yr ~ b1 * SWTSX  + b2 * VBMFX


```{r best fit}
#document of the package here: https://cran.r-project.org/web/packages/quadprog/quadprog.pdf
y_actual = matrix(alabama$InvestmentReturn_1yr)

x1 = matrix(alabama$SWTSX)
x2 = matrix(alabama$VBMFX)
x = cbind(x1, x2)

Dmat <- solve(chol(t(x) %*% x)) # matrix to be minimized
Amat <- cbind(rep(1,2), diag(2)) # vector define constraint
bvec <- c(1,rep(0,2)) #vector of b
dvec <- t(y_actual) %*% x  # vector to be minimized

result <- solve.QP(Dmat = Dmat, factorized = TRUE, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)


alabama_portfolio_result <- alabama %>% 
  add_column(b1_SWTSX = result$solution[1]) %>% 
  add_column(b2_VBMFX = result$solution[2]) %>% 
  mutate(portfolio_return = SWTSX*b1_SWTSX + VBMFX*b2_VBMFX) %>% 
  
  # sum of square total: (actual value - avarage of actual value)^2
  mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 

  # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
  mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 

  # r_squared
  mutate(r_quared = 1 - (sse/sst))

alabama_portfolio_result

```
## Comparison Result 

```{r comparison result}
# I can't believe R does not have a buit-in function for this. 

geo_mean <- function(vector) {
  vector <- vector + 1
  r <- prod(vector)^(1/length(vector)) - 1
  return(r)
}

comparison_result <- alabama_portfolio_result %>% 
  #geomean
  mutate(geomean_actual = geo_mean(InvestmentReturn_1yr)) %>% 
  mutate(geomean_portfolio = geo_mean(portfolio_return)) %>% 
  
  # standard deviation (i.e. risk)
  mutate(sd_actual = sd(InvestmentReturn_1yr)) %>% 
  mutate(sd_portfolio = sd(portfolio_return))

comparison_result
```

# Making a Function to Scale Up

```{r pension_benchmark_comparison function}
pension_benchmark_comparison <- function (planname) {
  plan <- pension_index %>% filter(PlanName == planname)
  
# optimization
  y_actual = matrix(plan$InvestmentReturn_1yr)

  x1 = matrix(plan$SWTSX)
  x2 = matrix(plan$VBMFX)
  x = cbind(x1, x2)
  
  Dmat <- solve(chol(t(x) %*% x)) # matrix to be minimized
  Amat <- cbind(rep(1,2), diag(2)) # vector define constraint
  bvec <- c(1,rep(0,2)) #vector of b
  dvec <- t(y_actual) %*% x  # vector to be minimized
  
  result <- solve.QP(Dmat = Dmat, factorized = TRUE, dvec = dvec, Amat = Amat, bvec = bvec, meq = 1)
  
  plan_result <- plan %>% 
    add_column(b1_SWTSX = result$solution[1]) %>% 
    add_column(b2_VBMFX = result$solution[2]) %>% 
    mutate(portfolio_return = SWTSX*b1_SWTSX + VBMFX*b2_VBMFX) %>% 
    
    # sum of square total: (actual value - avarage of actual value)^2
    mutate(sst = sum((InvestmentReturn_1yr - mean(InvestmentReturn_1yr))^2)) %>% 
  
    # sum of square error: (actual value - predicted value (i.e. portfolio return))^2
    mutate(sse = sum((InvestmentReturn_1yr - portfolio_return)^2)) %>% 
  
    # r_squared
    mutate(r_quared = 1 - (sse/sst))
  
  comparison_result <- plan_result %>% 
  #geomean
  mutate(geomean_actual = geo_mean(InvestmentReturn_1yr)) %>% 
  mutate(geomean_portfolio = geo_mean(portfolio_return)) %>% 
  
  # standard deviation (i.e. risk)
  mutate(sd_actual = sd(InvestmentReturn_1yr)) %>% 
  mutate(sd_portfolio = sd(portfolio_return)) %>% 
   
  
  # standard deviation (i.e. risk)
  mutate(sd_actual = sd(InvestmentReturn_1yr)) %>% 
  mutate(sd_portfolio = sd(portfolio_return))
  return(comparison_result) 
}

```
# Function Applied: California PERF
```{r California}
pension_benchmark_comparison("California PERF")
```

# Function Applied: South Dakota RS
```{r South Dakota RS}
pension_benchmark_comparison("South Dakota RS")
```

# Loop through all plans
```{r}
# Get a list of the first 5 plans
plan_list <- (unique(pension_index$PlanName))[1:5]

# Loop through each plan to produce result
for (plan in plan_list) {
  result <- pension_benchmark_comparison(plan)
  print(result)
}
  
```

